NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2019-02-18 Manuscript/manuscriptPS.tex

\section{Conclusions}
Lipids with PS headgroups, and their interactions with ions,
play an important role in lipid-mediated signaling
processes \cite{leventis10,yeung08}. Attemps to use MD
simulations to interpret the spectroscopic data have produced contradictory results for the calcium binding details to PS
headgroups \cite{melcrova16,valentine18,hallock18}.  
Here (as was previously done for PC lipids \cite{botan15,catte16})
we used the headgroup C--H bond order parameters and the open collaboration approach to evaluate the quality
of the headgroup structure and ion binding affinity to PS lipids
in available MD force fields.
The main advantage of this approach is the direct connection
between the accurately measured experimental order parameters and the simulations,
which reduces the ambiguity in the interpretation of experiments.

First, we complemented the available experimental data \cite{browning80,roux90} by measuring the signs of the PS headgroup order parameters.
Comparison to these data revealed that none of the available force fields
%tested using the NMRlipids open collaboration
reproduce the PS headgroup order parameters within the experimental accuracy. However,
the best models for the serine headgroup suggested a characteristic rigid conformation for its
carboxyl group. Comparison to the experimentally observed order parameters in POPC:POPS (5:1) bilayers  at varying ion
concentrations \cite{roux90} then showed that the tested MD force fields
overestimate the cation binding affinity to these bilayers with two exceptions. 1) The MacRog simulation with potassium counterion appear to produce the most realistic monovalent ion binding
affinity to PS-containing lipid bilayers. 2) The CHARMM36 force field with the recently introduced
NBfix correction for calcium \cite{kim16} underestimated the calcium binding affinity.
The experimentally measured response of the PS headgroup order parameter to the bound calcium, and to the dilution of bilayer with zwitterionic PC lipids, were not
qualitatively reproduced in any of the tested force fields, which underlines the need for more accurate
the MD force fields to study interactions between PS lipids and other biomolecules.
This is in contrast to the previous results with PC lipids,
where the experimentally measured headgroup order parameter responses to the bound charge
were in qualitative agreement even though the headgroup structures themselves were
incorrect and the cation binding affinities overestimated \cite{catte16}.

>>>>

\section{Conclusions}
Lipids with PS headgroups, and their interactions with ions,
play an important role in lipid-mediated signaling
processes \cite{leventis10,yeung08}. Recent studies using MD
simulations to interpret the various spectroscopic data
give contradictory results for the calcium binding details to PS
headgroups \cite{melcrova16,valentine18,hallock18}.  
Here, as was previously done for PC lipids \cite{botan15,catte16},
we used the headgroup C--H bond order parameters and the open collaboration approach to evaluate the quality
of the headgroup structure and ion binding affinity to PS lipids
in available MD force fields.
The main advantage of this approach is the direct connection
between the accurately measured experimental order parameters and the simulations,
which reduces the ambiguity in the interpretation of experiments.

First, we complemented the available experimental data of PS
lipid headgroup order parameters \cite{browning80,roux90} by measuring the signs of the order parameters.
Comparison to these data revealed that none of the available force fields
%tested using the NMRlipids open collaboration
was accurate enough to reproduce the PS
headgroup order parameters within the experimental accuracy. However,
the best models for the serine headgroup suggested a characteristic rigid conformation for its
carboxyl group. Comparison to the previously
measured headgroup order parameters from POPC:POPS (5:1) bilayers with different ion
concentrations \cite{roux90} then showed that the tested MD force fields
overestimate the cation binding affinity to the negatively charged bilayers
containing PS lipids with two exceptions. 1) The apparently most realistic monovalent ion binding
affinity to PS-containing lipid bilayers was observed in the MacRog
simulations with potassium counterions. 2) The CHARMM36 force field with the recently introduced
NBfix correction for calcium \cite{kim16} underestimated the calcium binding affinity.
The experimentally measured trends of the PS headgroup order parameter response
to the bound calcium, and to the dilution of bilayer with zwitterionic PC lipids, were not
qualitatively reproduced in any of the tested force fields, which indicates that improvements in
the MD force fields are necessary to study interactions between PS lipids and other biomolecules.
This is different to the previous results with PC lipids,
where the experimentally measured headgroup order parameter responses to the bound charge
were qualitative reproduced even though the headgroup structures themselves were
incorrect and the cation binding affinities overestimated \cite{catte16}.
